Concentration of Maxima and Fundamental Limits in High-Dimensional Testing and Inference

### 2021-02-03 5.URS/5.proofs-Ramsey.tex
DONE!
Ready for one final pass. Specifically, in the PDF, pls look at pages:

29, 30, 60 (notation <...>) Proofs of Propositons 5.1 and 5.3, Def. 5.3.

Page 115.

There is only one UNDEFINED reference on page 115. I can fix it once I understand the context better.

This result is a consequence of the celebrated work of \citet{ramsey2009problem}, which 
gave birth to Ramsey Theory (see e.g., \citet{conlon2015recent}).  
The Ramsey Theorem and the upper bound \eqref{eq:Ramsey-number} (established first in \cite{erdos1935combinatorial}) are at the heart of the proof of the following result.
%An excellent introduction to Ramsey theory is given in \url{http://math.mit.edu/~fox/MAT307-lecture05.pdf}.

@@

\begin{proof}[Proof of Proposition \ref{prop:lower-bound-correlation-Ramsey}]
By using \eqref{eq:Ramsey-number} and a refinement of the Stirling's formula, 
we will show at the end of the proof that for $k \le \log_2({n})/2$, we have 
\begin{equation}\label{eq:Ramsey-bounds}
 R(k,k) \le n,
\end{equation}
where $R(k,k)$ is the Ramsey number.

Now, construct a graph with vertices $\{1,\dots,n\}$ such that there is an edge between nodes $i$ and $j$ if and only if $\rho(i,j) > -\gamma$. 
In view of \eqref{eq:Ramsey-bounds} and Ramsey's theorem (see e.g., Theorem 1 in \cite{fox2009lecture} or \cite{conlon2015recent} for a recent survey on Ramsey theory), there is a subset of $k$ nodes $K =\{l_1,\dots,l_k\}$, which is either a {\em complete graph} or an {\em independent set}.

@@

\begin{proof}[Proof of Lemma \ref{lemma:positive-correlation}]
    Let $Z_1, \ldots, Z_{n+1}$ be random variables with covariance matrix $P$.
    Denote $\rho_j = \rho(1,j)$ and define 
    \begin{equation}
      R(j) = 
      \begin{cases}
        \frac{1}{\sqrt{1-\rho_j^2}}\left(Z(j) - \rho_j Z(1)\right), &\mbox{if } \rho_j<1,\\
        R^* &\mbox{if } \rho_j=1,
      \end{cases}
    \end{equation}
    where $R^*$ is an arbitrary zero-mean, unit-variance random variable.
    It is easy to see that $\var(R(j)) = 1$, and
    \begin{align*}
    \cov\left(Z(i), Z(j)\right) &= \cov\left(\rho_i Z(1) + \sqrt{1-\rho_i^2} R(i), \; \rho_j Z(1) + \sqrt{1-\rho_j^2} R(j)\right) \\
        &= \rho_i\rho_j + \sqrt{1-\rho_i^2}\sqrt{1-\rho_j^2} \;\cov\left(R(i), R(j)\right) \\
        &\ge c^2 + \min\left\{\cov\left(R(i), R(j)\right), 0\right\}.
    \end{align*}

    Therefore, Relation \eqref{eq:positive-correltation-lemma-conclusion} would hold if we can find a set of indices $K = \{l_1,\ldots,l_k\}$ such that $\cov\left(R(i),R(j)\right)>-c^2/2$ for all $i,j\in K$, where $k=|K|=\lfloor\log_2\sqrt{n}\rfloor$.
    This, however, follows from Proposition \ref{prop:lower-bound-correlation-Ramsey} applied to $\left(R(j)\right)_{j=2}^{n+1}$ with $\gamma = c^2/2$, provided that 
    $$
    k = \lfloor\log_2\sqrt{n}\rfloor \ge \lceil 2/c^2 \rceil + 1.
    $$
    The last inequality indeed follows form the assumption that $n \ge 2^{2\lceil2/c^2\rceil+4}$.
\end{proof}

>>>>

This result is a consequence of the celebrated work of \citet{ramsey2009problem}, which 
gave birth to Ramsey Theory \citep[see e.g.,][]{conlon2015recent}.  
The Ramsey Theorem and the upper bound \eqref{eq:Ramsey-number} \citep[established first in][]{erdos1935combinatorial} are at the heart of the proof of the following result.  A recent improvement on the upper bound is given by \cite{sah:2020}.
%An excellent introduction to Ramsey theory is given in \url{http://math.mit.edu/~fox/MAT307-lecture05.pdf}.

@@

\begin{proof}%[Proof of Proposition \ref{prop:lower-bound-correlation-Ramsey}]
By using \eqref{eq:Ramsey-number} and a refinement of the Stirling's formula, 
we will show at the end of the proof that for $k \le \log_2({n})/2$, we have 
\begin{equation}\label{eq:Ramsey-bounds}
 R(k,k) \le n,
\end{equation}
where $R(k,k)$ is the Ramsey number.

Now, construct a graph with vertices $\{1,\dots,n\}$ such that there is an edge between nodes $i$ and $j$ if and only if $\rho(i,j) \ge -\gamma$. 
In view of \eqref{eq:Ramsey-bounds} and Ramsey's theorem (see e.g., Theorem 1 in \cite{fox2009lecture} or \cite{conlon2015recent} for a recent survey on Ramsey theory), there is a subset of $k$ nodes $K =\{l_1,\dots,l_k\}$, which is either a {\em complete graph} or an {\em independent set}.  Recall that in a
complete graph, every two nodes are connected with an edge; while in independent sets, no two nodes are connected.

@@

\begin{proof}[Proof of Lemma \ref{lemma:positive-correlation}]
    Let $Z_1, \ldots, Z_{n+1}$ be random variables with covariance matrix $P$.
    Denote $\rho_j = \rho(1,j)$ and define 
    \begin{equation}
      R_j = 
      \begin{cases}
        \frac{1}{\sqrt{1-\rho_j^2}}\left(Z_j - \rho_j Z_1\right), &\mbox{if } \rho_j<1,\\
        R^* &\mbox{if } \rho_j=1,
      \end{cases}
    \end{equation}
    where $R^*$ is an arbitrary zero-mean, unit-variance random variable.
    It is easy to see that $\var(R_j) = 1$, and
    \begin{align*}
    \cov\left(Z_i, Z_j\right) &= \cov\left(\rho_i Z_1 + \sqrt{1-\rho_i^2} R_i, \; \rho_j Z_1 + \sqrt{1-\rho_j^2} R_j\right) \\
        &= \rho_i\rho_j + \sqrt{1-\rho_i^2}\sqrt{1-\rho_j^2} \;\cov\left(R_i, R_j\right) \\
        &> c^2 + \min\left\{\cov\left(R_i, R_j\right), 0\right\}.
    \end{align*}

    Therefore, Relation \eqref{eq:positive-correltation-lemma-conclusion} would hold if we can find a set of indices $K = \{l_1,\ldots,l_k\}$ such that $\cov\left(R_i,R_j\right)\ge -c^2/2$ for all $i,j\in K$, where $k=|K|=\lfloor\log_2\sqrt{n}\rfloor$.
    This, however, follows from Proposition \ref{prop:lower-bound-correlation-Ramsey} applied to $\left(R_j\right)_{j=2}^{n+1}$ with $\gamma = c^2/2$, provided that 
    $$
    k = \lfloor\log_2\sqrt{n}\rfloor \ge \lceil 2/c^2 \rceil + 1.
    $$
    The last inequality indeed follows form the assumption that $n \ge 2^{2\lceil2/c^2\rceil+4}$.
\end{proof}
